\center \fontsize{18.8}{20} \selectfont Towards Globally Color-Dual All-Loop Integrands

### 2023-09-20 notes/2loopNLSM.tex
some edits in intro & abstract

  We make progress towards understanding how color-dual
  representations must be constructed at the multi-loop level. First
  we identify a cubic theory, \emph{semi-abelian} Yang--Mills, that
  unifies many of the color-dual theories constructed in the
  literature, and provides a construction of color-dual numerators at
  one-loop directly from Feynman rules. Joint consideration of these
  theories highlights that the lifting of one-loop color-kinematics
  dual numerators to two loops may be just as challenging as the step
  from tree level to one loop.  As a first step in understanding the
  two-loop challenge, we construct color-dual two-loop four-point
  non-linear sigma model (NLSM) numerators using generalized unitarity
  methods.  The double-copy of the constructed NLSM numerators with
  themselves, $\mathcal{N}=0$ Yang--Mills, and $\mathcal{N}=4$
  Yang--Mills correctly reproduce the known physical unitarity cuts of
  special Gallileons, Born-Infeld theory, and
  Dirac-Born-Infeld-Volkov-Akulov theory, respectively.  These NLSM
  numerators are in explicit tension with generalizations of the
  one-loop NLSM numerators. \ace{I realize that we all know this, but
    does/will a statement to this effect actually make it into the
    draft?}
We find that the most general \emph{local} numerator for two-loop four-point numerator for $\mathcal{N}=0$ YM fails global color-kinematics duality and we pinpoint the failure to the kissing-triangles cut.

@@

\section{Introduction}\label{intro}
Since the turn of the century, our understanding of the $S$-matrix and
its hidden structures has expanded dramatically. At the heart of this
progress is the mantra that physical observables are simpler when
studied on-shell
\cite{Parke:1986gb,TasiLance,Cheung:2017pzi}. However, while plugging
in on-shell states certainly offers a path towards taming the
factorial growth of Feynman diagrams, it also obscures the off-shell
simplicity at the heart of many quantum field theories. A shining
example of this hidden structure is the duality between color and
kinematics \cite{BCJ,Bern:2010ue,BCJreview}, which states that the
kinematic numerators of gauge theory amplitudes can be rearranged to
obey the same algebraic identities as the constituent color
factors. When this duality is realized globally off-shell, it
dramatically decreases the combinatorial complexity posed by integrand
construction, and offers a path towards efficient assembly of quantum
gravity integrands directly from simpler gauge theory building blocks
via the double copy \cite{BCJ,Bern:2010ue}.

@@

Faced with this obstruction, we study the multi-loop sector of both
NLSM and Yang-Mills in \cref{2loopBoot} using an ansatz based
color-dual bootstrap. In contrast to much of the available literature
on construction of color-dual integrands at two-loop
\cite{Bern:2013yya,Mogull:2015adi,Johansson:2017bfl}, our results are
completely agnostic to the spacetime dimension; all the polarizations
and momenta appearing in our construction will remain formally
$D$-dimensional. This makes our methods and results particularly well
suited for algorithmically extracting rational terms from the
$D$-dependence of dimensionally regulated loop momenta. We are able to
achieve this by employing a custom solver, \texttt{FiniteFieldSolve},
which will soon be made public. \ace{This either needs an extra
  context sentence, or moved out of the intro.}

With this solver, we are able to compute a two-loop integrand for NLSM
that globally manifests the duality off-shell. This represents a new
benchmark for color-dual representations in non-supersymmetric gauge
theories.
However in sharp contrast, we find that Yang-Mills
does not permit a globally color-dual representation, even when considering the most general
polynomial ansatz of Lorentz covariant kinematics.\footnote{Globally color-dual local numerators for two-loop pure YM have been studied before but only in the context of a limited ansatz subject to certain loop power counting constraints \cite{Bern:2015ooa}.}
Concretely, the failure point can be pinpointed to a conflict between between the ``bowtie'' cut and the following Jacobi triple
\begin{equation}
   \NMCutC
  \Rightarrow
  \JacobiOneDoubleBox +  \JacobiOneCrossedBox + \JacobiOnePentaTriangle \neq 0 \,.
\end{equation}
We comment on the implications of this finding in \cref{conclusions},
and discuss alternative approaches where one must relax field
theoretic locality in order to accommodate the construction of
globally color-dual multi-loop integrands.

@@

\subsection{Color-dual Lagrangians}\label{offShellCK}
While the BCJ relations are an on-shell statement about the color-dual
nature of scattering amplitudes, one might aspire to hard code
color-kinematics duality by realizing an off-shell Lagrangian
description of the theory. The double-copy construction of
\cref{DCconstruction} would suggest a manifestly cubic Lagrangian of
the form
\begin{equation}
\mathcal{L} = P_V^{ab}P_W^{ij}\mathcal{O}^a_i \Box \mathcal{O}^b_j +  V^{abc} W^{ijk}\mathcal{O}^a_i \mathcal{O}^b_j \mathcal{O}^c_k
\end{equation}
where $V^{abc}$ and $W^{ijk}$ are cubic Feynman rules mixing field
operators $\mathcal{O}^{a}_i$ indexed by quantum numbers $a$ and
$i$. We have introduced $P_V^{ab}$ and $P_W^{ij}$ projection operators
to encode non-local kinetic structure that could in principle
participate in the construction.
The theory will be color dual if the cubic vertex is antisymmetric and satisfies the Jacobi identity, specifically,
\begin{align}
\textit{anti-symmetry}:& \quad{}_{a_1}\langle V^{a_2} \rangle_{a_3}+\text{cyc}(23)=0
\\
\textit{Jacobi identity}:& \quad {}_{a_1}\langle V^{a_2} V^{a_3}\rangle_{a_4}+\text{cyc}(234)=0
\end{align}
where 
\begin{equation}
{}_{a_1}\langle V^{a_2}V^{a_3}\cdots V^{a_{n-1}} \rangle_{a_n} \equiv V^{a_1a_2b_1}V^{b_1a_3b_2}\cdots V^{b_{n-3} a_{n-2}a_n}\,.
\end{equation}
Off-shell descriptions of
color-dual theories are rare, and only a few examples are available in
the literature \cite{Monteiro2011pc, Cheung:2016prv, Cheung:2021zvb,
  Cheung:2020djz, Ben-Shahar:2022ixa, Ben-Shahar:2021zww,
  Ben-Shahar:2021doh}.
Typically when one goes about constructing color-dual
theories directly from a set of cubic interactions, higher point
Jacobi and antisymmetry constraints require introducing additional
operators. See ref \cite{Ben-Shahar:2022ixa} for a recent
example. Indeed, as we will show in \cref{sec:cubic}, Yang-Mills
requires introducing a cubic 2-form interaction in order to manifest
the kinematic algebra just at four-point. Absent information about the
kinematic algebra, one can make progress in the construction of
loop-level color-dual amplitudes by employing a color-dual bootstrap.

@@

When an off-shell color-dual Lagrangian is not known, the amplitude integrand must be constructed from an ansatz.  The four conditions we impose on an integrand ansatz are summarized
here and elaborated on below.\footnote{The four ansatz conditions are
  similar to those of Ref.~\cite{Bern:2015ooa} except that there the
  authors demand that the numerators have the same loop power counting
  as Feynman gauge Feynman rules.  This is an additional assumption
  that is most emphatically not made here.}
\begin{enumerate}
\item \textbf{Off-shell Locality}: Diagram numerators are polynomials
  in momenta and polarizations, e.g., the diagrams only have poles given
  by the propagators of the diagrams.
\item \textbf{Color-kinematics duality}: The numerators obey the same
  Jacobi identities as the color factors.  This implicitly requires
  that the integrand is expressed in terms of purely cubic diagrams.
\item \textbf{Graph symmetries}: The diagram numerators are only a
  function of the diagram's topology and labeling.  Furthermore, the
  diagram numerators are invariant under the automorphisms of the
  diagrams including signs that compensate for color-factor sign
    changes.
\item \textbf{On-shell Unitarity }: The cuts of the ansatz must
  reproduce the physical unitarity cuts of the theory when internal
  momenta are taken on-shell.
\end{enumerate}

>>>>

  We make progress towards understanding how color-dual
  representations must be constructed at the multi-loop level. First
  we identify a cubic theory, semi-abelian Yang--Mills, that
  unifies many of the color-dual theories studied in the
  literature, and provides a prescriptive approach for constructing color-dual numerators at
  one-loop directly from Feynman rules. Consideration of this
  theory highlights that the lifting from one-loop color-kinematics
  dual numerators to two loops is just as challenging as the progression
  from tree-level to one-loop.  As a first step in understanding the
  two-loop challenge, we construct color-dual two-loop four-point
  nonlinear sigma model (NLSM) numerators using a $D$-dimensional color-dual bootstrap.  The double-copy of the constructed NLSM numerators with
  themselves, pure Yang--Mills, and $\mathcal{N}=4$
  super-Yang-Mills correctly reproduce the known physical unitarity cuts of
  special Gallileons, Born-Infeld theory, and
  Dirac-Born-Infeld-Volkov-Akulov theory, respectively.  These NLSM
  numerators are in explicit tension with generalizations of the
  one-loop NLSM numerators. Using our constructive bootstrap, we also find that the most general \emph{local} two-loop four-point numerator for pure Yang-Mills fails to satisfy global color-kinematics duality. We pinpoint the failure to the bow-tie unitarity cut, and discuss a path towards \emph{non-local} construction of color-dual integrands at generic loop order.

@@

\section{Introduction}\label{intro}
Since the turn of the century, our understanding of the $S$-matrix and
its concealed structures has expanded dramatically. At the heart of this
progress is the mantra that physical observables are simpler when
studied on-shell
\cite{Parke:1986gb,TasiLance,Cheung:2017pzi}. However, while plugging
in on-shell states certainly offers a path towards taming the
factorial growth of Feynman diagrams, it also obscures the off-shell
simplicity at the heart of many quantum field theories. A shining
example of this hidden structure is the duality between color and
kinematics \cite{BCJ,Bern:2010ue,BCJreview}, which states that the
kinematic numerators of gauge theory amplitudes can be rearranged to
obey the same algebraic identities as the constituent color
factors. When this duality is realized globally off-shell, it
dramatically decreases the combinatorial complexity posed by integrand
construction, and offers a path towards efficient assembly of quantum
gravity integrands directly from simpler gauge theory building blocks
via the double copy \cite{BCJ,Bern:2010ue}.

@@

Faced with this obstruction, we study the multi-loop sector of both
NLSM and Yang-Mills in \cref{2loopBoot} using an ansatz based
color-dual bootstrap. In contrast to much of the available literature
on color-dual integrand construction at two-loop
\cite{Bern:2013yya,Mogull:2015adi,Johansson:2017bfl}, our results are
completely agnostic to the spacetime dimension; all the polarizations
and momenta appearing in our construction will remain formally
$D$-dimensional. This makes our methods and results particularly well
suited for algorithmically extracting rational terms from the
$D$-dependence of dimensionally regulated loop momenta. Prior to this work, one of the primary barriers in studying multi-loop color-dual integrands $D$-dimensionally is the explosion of terms that one encounters in generic ansatz-based constructions at high-loop order. We were able overcome this barrier by employing a custom solver, \texttt{FiniteFieldSolve},
which will soon be made public. \ace{This either needs an extra
  context sentence, or moved out of the intro.}

With this solver, we are able to compute a two-loop integrand for NLSM
that globally manifests the duality off-shell. This represents a new
benchmark for color-dual representations in non-supersymmetric gauge
theories.
However in sharp contrast, we find that Yang-Mills
does \emph{not} permit a globally color-dual representation, even when considering the most general
polynomial ansatz of Lorentz covariant kinematics.\footnote{Globally color-dual local numerators for two-loop pure YM have been studied before but only in the context of a limited ansatz subject to certain loop power counting constraints \cite{Bern:2015ooa}.}
Concretely, the failure point can be pinpointed to a conflict between between the ``bowtie'' cut and the following Jacobi triple:
\begin{equation}
   \NMCutC
  \Rightarrow
  \JacobiOneDoubleBox +  \JacobiOneCrossedBox + \JacobiOnePentaTriangle \neq 0 \,.
\end{equation}
We comment on the implications of this finding in \cref{conclusions},
and discuss alternative approaches where one must relax field
theoretic locality in order to accommodate the construction of
globally color-dual integrands at generic loop order.

@@

\subsection{Color-dual Lagrangians}\label{offShellCK}
While the BCJ relations are an on-shell statement about the color-dual
nature of scattering amplitudes, one might aspire to hard code
color-kinematics duality by realizing an off-shell Lagrangian
description of the theory. The double-copy construction of
\cref{DCconstruction} would suggest a manifestly cubic Lagrangian of
the form
\begin{equation}
\mathcal{L} = P_V^{ab}P_W^{ij}\mathcal{O}^a_i \Box \mathcal{O}^b_j +  V^{abc} W^{ijk}\mathcal{O}^a_i \mathcal{O}^b_j \mathcal{O}^c_k
\end{equation}
where $V^{abc}$ and $W^{ijk}$ are cubic Feynman rules mixing field
operators $\mathcal{O}^{a}_i$ indexed by quantum numbers $a$ and
$i$. We have introduced $P_V^{ab}$ and $P_W^{ij}$ projection operators
to encode non-local kinetic structure that could in principle
participate in the construction.
The theory will be color dual if the cubic vertex is antisymmetric and satisfies the Jacobi identity, specifically,
\begin{align}
\textit{anti-symmetry}:& \quad{}_{a_1}\langle V^{a_2} \rangle_{a_3}+\text{cyc}(23)=0
\\
\textit{Jacobi identity}:& \quad {}_{a_1}\langle V^{a_2} V^{a_3}\rangle_{a_4}+\text{cyc}(234)=0
\end{align}
where 
\begin{equation}
{}_{a_1}\langle V^{a_2}V^{a_3}\cdots V^{a_{n-1}} \rangle_{a_n} \equiv V^{a_1a_2b_1}P_{b_1 c_1}V^{c_1 a_3b_2}\cdots P_{b_{n-3}c_{n-3}}V^{c_{n-3} a_{n-1}a_n}\,.
\end{equation}
Off-shell descriptions of
color-dual theories are rare, and only a few examples are available in
the literature \cite{Monteiro2011pc, Cheung:2016prv, Cheung:2021zvb,
  Cheung:2020djz, Ben-Shahar:2022ixa, Ben-Shahar:2021zww,
  Ben-Shahar:2021doh}.
Typically when one goes about constructing color-dual
theories directly from a set of cubic interactions, higher point
Jacobi and antisymmetry constraints require introducing additional
operators. See ref \cite{Ben-Shahar:2022ixa} for a recent
example. Indeed, as we will show in \cref{sec:cubic}, Yang-Mills
requires introducing a cubic 2-form interaction in order to manifest
the kinematic algebra just at four-point. Absent information about the
kinematic algebra, one can make progress in the construction of
loop-level color-dual amplitudes by employing a color-dual bootstrap.

@@

When an off-shell color-dual Lagrangian is not known, the loop-level amplitude integrand must be constructed from an ansatz.  The four conditions we impose on an integrand ansatz are summarized
here and elaborated on below.\footnote{The four ansatz conditions are
  similar to those of Ref.~\cite{Bern:2015ooa} except that there the
  authors demand that the numerators have the same loop power counting
  as Feynman gauge Feynman rules.  This is an additional assumption
  that is most emphatically not made here.}
\begin{enumerate}
\item \textbf{Off-shell Locality}: Diagram numerators are polynomials
  in momenta and polarizations, e.g., the diagrams only have poles given
  by the propagators of the diagrams.
\item \textbf{Color-kinematics duality}: The numerators obey the same
  Jacobi identities as the color factors.  This implicitly requires
  that the integrand is expressed in terms of purely cubic diagrams.
\item \textbf{Graph symmetries}: The diagram numerators are only a
  function of the diagram's topology and labeling.  Furthermore, the
  diagram numerators are invariant under the automorphisms of the
  diagrams including signs that compensate for color-factor sign
    changes.
\item \textbf{On-shell Unitarity }: The cuts of the ansatz must
  reproduce the physical unitarity cuts of the theory when internal
  momenta are taken on-shell.
\end{enumerate}
